---
title: "Exploratory Data Analysis"
author: "kbr1218"
format: html
editor: visual
---

```{r, setup, include=FALSE}
library(tidyverse)
library(Lahman)

team <- Teams %>% 
  filter(yearID >= 2000, yearID != 2020)

```

# Exploratory Data Analysis:

EDA: Visualize and transform data to systematically explore

1.  Generate Questions

2.  Search for Answers

    1.  Visualize

    2.  Transform

    3.  Model

3.  Use results to generate new questions

**KEY: All about asking GOOD questions -\> direct you to a part of your data**

Standard questions to start

1.  Question 1: What is the variation in my variables?

2.  Question 2: How are variables related? (co-variation)

# Tendency

#### Mean: average value

$$
\rm mean = \bar{x} = 
\frac{\displaystyle\sum\limits_{i=1}^{N}x_{i}}{N}
$$

i -\> individual observation

N -\> total number of observations

Can be affected by outliers

-   Trimmed mean: Drops values at each end, eliminate influence of outlier

-   Weighted mean: weight observation differently

#### Median: Middle number in sorted list

Robust to outliers

```{r, median, warning=FALSE, message=FALSE}


# Create Slug%

team %>% 
  summarise_if(is.numeric, list(mean = mean, median = median))
```

# Variability

variability: Dispersion of data values

### Measure: Deviations from central tendency

-   Deviation(편차)

-   standard deviation(표준 편차)

#### Variance: Average squared deviation (in squared units)

$$
\rm variance = \sigma^2 = \frac{\displaystyle\sum\limits_{i=1}^{N}
\left(x_{i}-\bar{x}
\right)^2}{N}
$$

#### Standard Deviation: square root of variance

#### Percentile : Pth percentile is value such that, at least P% of values take that value or less.

```{r, variation, message=FALSE, warning=FALSE}

```

### Visualize

ggplot(data = df) +

-   %\>% ggplot에서는 +

-   tidyverse 안에 ggplot이 있음

geom\_\<function\>(aes(\<mapping\>))

#### Boxplot: Median and Quartiles

```{r, boxplot, warning=FALSE, message=FALSE}

ggplot(data = team) + 
  geom_boxplot(aes(y = HR)) +           # 박스 모양의 플롯
  theme_bw()
```

#### Histogram: Count number of observations within a range

```{r, histogram, warning=FALSE, message=FALSE}

ggplot(data = team) + 
  geom_histogram(aes(x = HR), bins = 100) + 
  # bin이 많을 수록 x축의 range를 세분화시킴
  theme_light()
```

#### Density

```{r, density, warning=FALSE, message=FALSE}
ggplot(data = team) +
  geom_density(aes(x = HR)) + 
  theme_bw()

# histogram과 density를 같은 그래프에 넣기
ggplot(data = team, aes(x = HR)) + 
  geom_histogram(aes(y = ..density..)) +
  geom_density(col = "red") + 
  theme_bw()
```

# Relation

how are variables related (closely positive or negative)

### Measure

#### Correlation: Measure between -1 and 1.

왜 우리는 카운팅보다 비율(레이팅)을 더 선호할까?

$$
\rm correlation = r = \frac{\displaystyle\sum\limits_{i=1}^{N}
\left(x_{i}-\bar{x}\right)\left(y_{i}-\bar{y}\right)}
{\left(N-1\right)\sigma_x\sigma_y}
$$

$$
-1\leq r \leq 1
$$

```{r, relation, message=false, warning=FALSE}


cor(team$HR, team$R)
cor(team$SB, team$R)
cor(team$ERA, team$W)

```

### Visualize

#### Scatterplot

```{r, scatterplot, warning=FALSE, message=FALSE}
team <- team %>% 
  mutate(
    WPct = W/G
  )
  
ggplot(data = team) +
  geom_point(aes(x = HR, y = R,
                 col = WSWin,
                 size = WPct),
             alpha = 0.4) +
  theme_bw()
```

#### Corrplot

```{r, corrplot, warning=FALSE, message=FALSE}
library(corrplot)


corr_vars <- team %>% 
  select(HR, R, SO, ERA, W)

corr_matrix <- round(cor(corr_vars), digits = 2)

corrplot(corr_matrix, method = "number")
```

# Exercises

Use FanGraphs or Baseball Savant to build a data set and:

<https://baseballsavant.mlb.com/>

-   -\> leaderboard -\> csv 파일 다운받고 rename

1.  Explore an individual pitching metric, analyze a visualization you created, and use that visualization to generate an interesting question.

2.  Explore an individual hitting metric, analyze a visualization you created, and use that visualization to generate an interesting question.

3.  Explore a team pitching metric, analyze a visualization you created, and use that visualization to generate an interesting question.

4.  Explore a team hitting metric, analyze a visualization you created, and use that visualization to generate an interesting question.
