---
title: "Baseball Metrics"
author: "kbr1218"
format: html
editor: visual
---

# Basics (Phase 1)

#### Batting Average (BA): Hits per at bat

$$
BA=\frac{H}{AB}
$$

Issues: Good plate appearance outcomes, that don't count

-   Plate appearance outcomes:

    1.  Hit (single, double, triple, home run)

    2.  **Walk** (3 strike // 4 balls) -\> Not in AB

    3.  Generic Out (K, Ground out, fly out)

    4.  Sacrifice Bunt -\> Not in AB

    5.  Sacrifice Fly -\> Not in AB

    6.  Hit by pitch -\> Not in AB

    7.  Fielders Choice

    8.  Reach on Error

    9.  Dropped 3rd Strike

    10. Defensive Interference -\> Not in AB

        -\> 이 10개 중 **(1, 3, 7, 8, 9)**만 bat으로 count됨

-   Worse: Not all hits are equal (not all contact is equal)

    -   ex\> single과 home run 모두를 batting average는 1번이라고 생각함 (둘 사이에는 큰 차이가 있음에도 불구하고)

#### Run Batted In (RBI) : Result of a plate appearance is a run being scored (not if because of error or double play)

|          | PA  | RBI | PA ROB | %PA ROB | PA RISP | %PA RISP | BA w/ ROB |
|----------|-----|-----|--------|---------|---------|----------|-----------|
| Player A | 636 | 101 | 315    | 49.5    | 195     | 30.7     | .247      |
| Player B | 608 | 79  | 231    | 38      | 130     | 21.4     | .325      |

Issue: Product of opportunity (B의 BA가 더 높지만, A가 기회가 더 많아서 높은 RBI를 가지고 있음)

-\> Count metric (커런시의 수를 세는 것) versus a rate metric (비율)

-\> RBI / PA ROB를 해서 비교하면 player B가 더 높음

-\> RBI / %PA ROB를 해서 비교하면 player B가 더 높음

-\> B가 더 좋은 선수

#### Error (E): Official scorer judgement if player fails to convert an out on a play an average fielder should have make. Also, poor play that allows runner to advance

[Error Example](https://baseballsavant.mlb.com/sporty-videos?playId=1f7d22d8-b80a-4281-bb73-a78f88489253)

[Hit Example](https://baseballsavant.mlb.com/sporty-videos?playId=0cca465e-5311-4710-84dd-f879c2f80cd4)

Issue: subjective and doesn't account for range doesn't factor mental mistakes

-   만약 내가 잘 받는게 아니라 내가 위치를 잘 정하는 거라면

-   error가 많을 수록 Fielder가 적어짐 -\> 이걸 배경으로 돈을 받는데 객관적이지 않음 subjective함

**Pitcher Win (W): Look up on MLB.com**

Issue: Don't have to pitch effectively to get a win, can be subjective, can reward ineffective pitcher, depends on offense and defense

#### Save (SV): **Look up on MLB.com**

Issue: Don't have to be effective, not necessarily most important situation

# Better Metrics (Phase 2)

**What makes a good metric?**

1.  Measure contribution to success

2.  Repeatable (signal)

    noise랑 구분됨 (noise는 랜덤으로 나타나는 것)

#### On-Base Percentage (OBP): How frequently hitter reaches base (doesn't make an out) $$
OBP=\frac{H+BB+HBP}{AB+BB+HBP+SF}
$$

1.  Strength: More complete measure of a hitter's contribution to success, outcome -\> skill

2.  Weakness: Weights all ways of getting on base the same, on base due to error doesn't count (hit hard and fast)

#### Slugging Percentage (SLG): Average bases per at bat

$$
SLG=\frac{1\times1B+2\times2B+3\times3B+4\times HR}{AB}
$$

1.  Strength: Values extra base hits more heavily, "power" rate metric

2.  Weakness: leaves out a lot of place appearances (walks), "weights" are not necessarily accurate measure of production

#### On-Base Plus Slugging (OPS): Ability to get on base and hit for power

$$
OPS = OBP + SLG
$$

1.  Strength: most correlated to runs

2.  Weakness: add different denominators (not interpret-able), doesn't account for hitting environment

```{r, Create Metrics, echo=FALSE, message=FALSE, warning=FALSE}

library(Lahman)
library(tidyverse)
library(kableExtra)
library(patchwork)

team <- Teams
batting <- Batting
players <- People %>% 
  mutate(
    Name = paste(nameFirst, nameLast, sep = " ")
  )

# Create BA, OBP, SLG, OPS for team and individual
team_2000 <-team_2000 %>% 
  mutate(
    X1B = H - X2B - X3B - HR, 
    BA = round(H/AB, digits = 3), 
    PA = AB + BB + HBP + SF, 
    OBP = round((H + BB + HBP)/PA, digits = 3), 
    SLG = round((1 * X1B + 2 * X2B + 3 * X3B + 4 * HR)/AB, digits = 3), 
    OPS = OBP + SLG
  )

# Team Level
team_2000 %>% 
  select(franchID, yearID, OBP, SLG, BA) %>% 
  arrange(-OBP) %>% 
  head(10) %>% 
  kable(booktabs = T) %>%                # making table with kable
  kable_styling(bootstrap_option = c('striped'))


# Player Level


# OPS Leaderboard


# Percentile

```

##### Relate to Success

What wins games? Score more runs than opponent. For offense objective is to score runs. Measure success must see how events lead to runs.

// offensive event -\> runs -\> wins

```{r, success, message=FALSE, warning=FALSE}

team_2000 %>% 
  summarise (
    r_BA = cor(BA, R), 
    r_OBP = cor(OBP, R),     # 높은 OBP을 가지고 있는 팀이 점수를 더 많이 얻음
    r_SLG = cor(SLG, R),     # 높은 SLG을 가지고 있는 팀이 점수를 더 많이 얻음
    r_OPS = cor(OPS, R)      # 이 중에서 correlation이 가장 높음
  ) %>% 
  kable(booktabs = T) %>% 
  kable_styling()

# 팀마다 각각의 BA/OBP/SLG/OPS가 R(득점)과 얼마나 연관이 되어있는지 시각화

p1 <- ggplot(data = team_2000, aes(x = BA, y = R)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_bw()

p2 <- ggplot(data = team_2000, aes(x = OBP, y = R)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_bw()

p3 <- ggplot(data = team_2000, aes(x = SLG, y = R)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_bw()

p4 <- ggplot(data = team_2000, aes(x = OPS, y = R)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_bw()


(p1|p2)/(p3|p4) # 2x2로 그래프 삽입
```

##### Repeatable

Year-to-year consistency? Signal or noise. Pay for signal, not for noise.

```{r, repeatable, message=FALSE, warning=FALSE}
bat_2000 <- batting %>%         # 팀 단위가 아닌 개인 선수 기준으로 분석
  filter(yearID >= 2000, yearID != 2022) %>% 
  group_by(playerID, yearID) %>% 
  summarise_if(is.numeric, sum) %>% 
  filter(AB > 100)

bat_2000 <- bat_2000 %>%       # season stat
  mutate(
    X1B = H - X2B - X3B - HR, 
    BA = round(H/AB, digits = 3), 
    PA = AB + BB + HBP + SF, 
    OBP = round((H + BB + HBP)/PA, digits = 3), 
    SLG = round((1 * X1B + 2 * X2B + 3 * X3B + 4 * HR)/AB, digits = 3), 
    OPS = OBP + SLG
  )

bat_2000 <- bat_2000 %>% 
  group_by(playerID) %>% 
  arrange(yearID) %>% 
  mutate(
    BA_lag = lag(BA),            # 2002라면 2001 데이터를 줌 (없으면 NA)
    OBP_lag = lag(OBP),
    SLG_lag = lag(SLG), 
    OPS_lag = lag(OPS)
  ) %>% 
  ungroup()     # ungroup하지 않으면 플레이어 단위로 correlation을 찾으려고 할거임

# 이것들이 correlated되어있는지 확인
bat_2000 %>% 
  summarise(
    BA_rept = cor(BA, BA_lag, use = "complete.obs"),
    # BA는 별로 반복적이지 않음 (운이 따라야 하는 경우가 있기 때문)
    OBP_rept = cor(OBP, OBP_lag, use = "complete.obs"), 
    SLG_rept = cor(SLG, SLG_lag, use = "complete.obs"), 
    OPS_rept = cor(OPS, OPS_lag, use = "complete.obs")
  ) %>% 
  kable(booktabs = T) %>% 
  kable_styling()

ggplot(data = bat_2000, aes(x = BA_lag, y = BA)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, col = "red", linetype = "dashed") +    # 붉은 선은 45도
  theme_bw()

```

# Exercises

1.  Use a correlation metric and show graphically whether the following metrics are repeatable year-to-year. Use pitchers with 30 innings pitched or more (IP = IPouts/3). Note: we're going to use rate metrics instead of count.

    1.  RBI per Plate Appearance

    2.  Win per Pitching Appearance

    3.  Earned run average (ERA)

    4.  Walk and Hits Per Inning Pitched (WHIP)

2.  Use a correlation metric and show graphically which of the following pitching stats is most related to the measure of success Runs Allowed. Note: This depends on the defense so it's not perfect!

    1.  Win per Pitching Appearance

    2.  Earned run average (ERA)

    3.  Walk and Hits Per Inning Pitched (WHIP)

    4.  Strikeouts (K) per Inning Pitched (IP)

    5.  Walks (BB) per Inning Pitched
